---
title: "Exercise on Diamonds data"
author: "Nicole Li"
date: "3/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setting}
library(ggplot2)
library(dplyr)

```
<br/>
```{r preview data}
glimpse(diamonds)
```
<br/>
```{r}
summary(diamonds$carat)
quantile(diamonds$carat, 0.98)
```
<br/><br/>

```{r}
mydiamonds <- diamonds[which(diamonds$carat<=2.04),]
```
<br/><br/>


##### Chanllenge 1  
The diamonds data.frame is included in the ggplot2 package. Study the relationships between carat, price and color. 
```{r}
ggplot(mydiamonds, aes(x=carat, y=price, color=color)) +
  theme(legend.position = 'right') +
  geom_point(size=0.5) +
  geom_smooth(method='loess', se=FALSE, fill=NA, size=0.5) +
  scale_x_continuous(name="Carat", limits=c(0,2.1), breaks=seq(0,2.1,0.3)) +
  scale_y_continuous(name="Price") +
  ggtitle("Relationships Between carat, price And color") +
  theme_bw()
  
```
<br/><br/>


##### Chanllenge 2

Dive into the relationship between color and carat.   
```{r}
ggplot(diamonds, aes(x=color, y=carat)) +
  theme(legend.text=element_text(size=3)) +
  geom_boxplot(aes(fill=color), alpha=0.7, outlier.alpha=0.3) +
  ggtitle("Relationships Between carat And color") +
  theme_bw()
```
<br/>

```{r, fig.width=10, fig.height=3}
ggplot(diamonds, aes(x=carat, color=color)) +
  geom_density() +
  scale_x_continuous(limits = c(0, 3)) +
  facet_wrap(~color, ncol=10) +
  theme_bw()

```
<br/><br/>



##### Chanllenge 3   
Dive into the relationship between color and price.   
```{r}
ggplot(diamonds, aes(x=color, y=price)) +
  geom_boxplot(aes(color=color))
```


Density plots on price by color...
```{r}
ggplot(diamonds, aes(x=price)) +
  geom_density(aes(color=color)) +
  scale_x_continuous(limits = c(0, quantile(diamonds$price, 0.98))) +
  facet_wrap(~color, nrow = 2) +
  theme_bw()
```
<br/><br/>


##### Chanllenge 4     
Dive into the relationship between color and (price/carat).   
```{r}
ggplot(diamonds, aes(x=color, y=price/carat)) +
  geom_boxplot(aes(color=color), outlier.alpha = 0.5)
```
<br/>

Density plot
```{r, fig.width=10, fig.height=3}
ggplot(diamonds, aes(x=price/carat)) +
  geom_density(aes(color=color)) +
  scale_x_continuous(limits = c(0,12000)) +
  facet_wrap(~color, nrow=2)
```



