---
title: "Tutorial on ggplot"
author: "Nicole Li"
date: "3/12/2018"
output: 
  html_document:
    fig_height: 6
    fig_width: 8
---

# http://tutorials.iq.harvard.edu/R/Rgraphics/Rgraphics.html#exercise_iii

# `ggplot`
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(ggplot2)
housing <- read.csv("./dataSets/landdata-states.csv")
head(housing, 5)
```
<br/>

### `ggplot` VS. Basic Plots
* is more verbose for simple graphics
* is less verbose for complex graphics
* only for data.frame
* use a different system for adding plot elements
<br/>
1. For simple graphic
```{r}
# With base graph
hist(housing$Home.Value)
```

```{r}
# With ggplot
ggplot(housing, aes(x = Home.Value)) +
  geom_histogram()
```
<br/>

2. For complex graphic
```{r}
# With base graph
plot(Home.Value ~ Date, data = subset(housing, State=='MA'))
points(Home.Value ~ Date, data = subset(housing, State=='TX'), col='red')
legend(1975, 400000, title = 'State',
       c('MA','TX'), col=c('black','red'), pch=c(1,1))
```

```{r}
# With ggplot
ggplot(data = subset(housing, State %in% c('MA','TX')),
       aes(x=Date, y=Home.Value, color=State)) +
  geom_point()
```

<br/><br/>

### Aesthetics and Geometric Objects
*A. Aesthetic Mapping*: 
In ggplot, the "aesthetic" means "something you can see". For example:
  + position (e.g. on the x and y axis)
  + color (outside color)
  + fill (inside color)
  + shape
  + linetype
  + size

<br/>

*B. Geometric Objects*: 
These are the actual marks that we put onto a plot. For example:
  + points
  + lines
  + boxplot...
A plot must have at least one `geom` object, with no upper limit. 
```{r results='hide'}
help.search('geom_', package='ggplot2')
```
<br/>

1. Scatter/Point Plot
```{r}
house2001Q1 <- subset(housing, Date==2001.25)
ggplot(house2001Q1, aes(x=Land.Value, y=Structure.Cost)) + # mapping for x and y
  geom_point()
```
<br/>

2. Lines (Prediction Lines)
```{r}
house2001Q1$pred_sc <- predict(lm(Structure.Cost ~ log(Land.Value), house2001Q1))

ggplot(house2001Q1, aes(x=log(Land.Value), y=Structure.Cost)) +
  geom_point(aes(color=Home.Value)) +
  geom_line(aes(y=pred_sc))
```
<br/>

3. Smoothers
```{r}
ggplot(house2001Q1, aes(x=log(Land.Value), y=Structure.Cost)) +
  geom_point(aes(color=Home.Value), size=1) +  # !!! size, the fixed aes, is set outside the "aes".
  geom_line(aes(y=pred_sc)) +
  geom_smooth()
```
<br/>

4. Text (Lable points)
```{r}
ggplot(house2001Q1, aes(x=log(Land.Value), y=Structure.Cost)) +
  geom_point(aes(color=Home.Value, shape=region)) +          # !!! mapping with other fields
  geom_text(aes(label=State), size=2, hjust=0.5, vjust=-0.9) # !!! size & hjust & vjust, the fixed aes, are set outside the "aes".
```
<br/>

### Exercise I
```{r}
# Human Development Index (HDI) & Corruption Perception Index (CPI)
dat <- read.csv("./dataSets/EconomistData.csv")
head(dat, 3)
```
<br/>

* Task 1: Create a scatter plot with CPI on the x axis and HDI on the y axis.
* Task 2: Color the points blue.
* Task 3: Map the color of the the points to Region.
* Task 4: Make the points bigger by setting size to 2
* Task 5: Map the size of the points to HDI.Rank
```{r}
ggplot(dat, aes(x=CPI, y=HDI)) +
  # geom_point(color='blue') +
  # geom_point(aes(color=Region), size=2)
  geom_point(aes(color=Region, size=HDI.Rank))
```
<br/><br/>




### Statistical Transformations
1. Statistical Transformations    
Some plots, like boxplot and prediction lines, require statistical transformation.
* for boxplot, the y values must be transformed to median and 1.5.
* for smoother, the y values must be transformed to the predicted values.
<br/>

2. Setting Statistical Transformation Arguments   
Arguments of `stat` functions can be passed through `geom_` functions. But in order to change it, you have to first determine which `stat` the `geom_` uses, then determine the arguments to that `stat`.
```{r}
ggplot(housing, aes(Home.Value)) +
  geom_histogram()
```
<br/>

To change the width of each bars in the histogram, we need to state `stat_bin` first...
```{r}
ggplot(housing, aes(Home.Value)) +
  geom_histogram(stat='bin', binwidth=4000)
```
<br/>


3. Changing The Statistical Transformation  

In `geom_bar`, the `stat` is `count` by default. 
```{r}
housing_agg <- aggregate(Home.Value~State + region, data=housing, FUN = mean)

ggplot(housing_agg, aes(x=State, y=Home.Value)) +
  geom_bar(aes(color=region), stat = 'identity') # specify the 'stat' value
```
<br/>


### Exercise II
* Re-create a scatter plot with CPI on the x axis and HDI on the y axis (as you did in the previous exercise).
* Overlay a smoothing line on top of the scatter plot using geom_smooth.
* Overlay a smoothing line on top of the scatter plot using geom_smooth, but use a linear model for the predictions. Hint: see ?stat_smooth.
* Overlay a smoothing line on top of the scatter plot using geom_line. Hint: change the statistical transformation.
* BONUS: Overlay a smoothing line on top of the scatter plot using the default loess method, but make it less smooth. Hint: see ?loess.
```{r}
ggplot(dat, aes(x=CPI, y=HDI)) +
  geom_point() +
  geom_smooth(method='loess', span=0.3) #  "lm", "glm", "gam", "loess", "rlm"
  # geom_line(stat='smooth', method='loess')
```
<br/><br/>


### Scales
1. Controlling Aesthetic Mapping
Aesthetic mapping (i.e., with aes()) only says that a variable should be mapped to an aesthetic. It doesn’t say how that should happen.  
Describing what colors/shapes/sizes etc. to use is done by modifying the corresponding `scale` (`scale_<aesthetic>_<type>`).   
The `scale` in ggplot2 includes:  
* position
* color and fill
* size
* shape
* line type
<br/>

2. Common Scale Arguments
```{r}
ggplot(housing, aes(x=State, y=Home.Price.Index)) +
  theme(legend.position = 'top', axis.text = element_text(size=6)) +
  geom_point(aes(color=Date), 
             alpha=0.5, 
             size=1, 
             position=position_jitter(width=0.15, height=0)) + # adding random noise to a plot to make it easier to read
  scale_x_discrete('State Abbreviation') +
  scale_color_continuous(name='Date', 
                         breaks=c(1976, 1994, 2010), labels=c("'76", "'94", "'10"),
                         low='blue', high='red') 
```
<br/>


3. Using different color scales    
ggplot2 has a wide variety of color scales.    
This is an example of using `scale_color_gradient2` to interpolate between three different colors.
```{r}
ggplot(housing, aes(x=State, y=Home.Price.Index)) +
  theme(legend.position = 'top', axis.text = element_text(size=5)) +
  geom_point(aes(color=Date), alpha=0.5, size=1, 
             position=position_jitter(width = 0.1, height = 0)) +
  
  scale_color_gradient2(name="",
                        breaks=c(1976, 1994, 2013),
                        labels=c("'76","'94","'13"),
                        low='blue', high='red', mid='grey60', midpoint = 1994)
```
<br/>  

##### Available Scales Categories:
* `scale_color_`
* `scale_fill_`
* `scale_size_`
* `scale_shape_`
* `scale_linetype_`
* `scale_x_`
* `scale_y_`

<br/><br/> 

### Exercise III   
* Create a scatter plot with CPI on the x axis and HDI on the y axis. Color the points to indicate region.
* Modify the x, y, and color scales so that they have more easily-understood names (e.g., spell out “Human development Index” instead of “HDI”).
* Modify the color scale to use specific values of your choosing. Hint: see ?scale_color_manual.
```{r}
ggplot(dat, aes(x=CPI, y=HDI)) +
  theme(text=element_text(size=9)) +
  geom_point(aes(color=Region)) +
  scale_x_continuous(name="Corruption Perception Index") +
  scale_y_continuous(name="Human Development Index") +
  scale_color_manual(name="Region of World", 
                     values=c("#24576D",
                              "#099DD7",
                              "#28AADC",
                              "#248E84",
                              "#F2583F",
                              "#96503F"))
```



### Facet
1. Faceting
```{r}
# Without faceting, all the lines are crowding together and it's hard to distinguish by state.
ggplot(housing, aes(x=Date, y=Home.Value)) +
  theme(text = element_text(size=8)) +
  geom_line(aes(color=State))
```


```{r}
# With faceting, all the 
ggplot(housing, aes(x=Date, y=Home.Value)) +
  theme(text = element_text(size=8)) +
  geom_line() +
  facet_wrap(~State, ncol=10)
```




### Other Examples

```{r}
library(tidyr)
housing.byyear <- aggregate(cbind(Home.Value, Land.Value) ~ Date, data=housing, mean)
home.land.byyear <- gather(housing.byyear, value = "value", key = "type", Home.Value, Land.Value)

ggplot(home.land.byyear, aes(x=Date, y=value, color=type)) +
  geom_line()

```


##### Graph Simulation
```{r}
library(ggthemes)
library(ggrepel)

dat <- read.csv("./dataSets/EconomistData.csv")
pointsToLabel <- c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan",
                   "Afghanistan", "Congo", "Greece", "Argentina", "Brazil",
                   "India", "Italy", "China", "South Africa", "Spane",
                   "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France",
                   "United States", "Germany", "Britain", "Barbados", "Norway", "Japan",
                   "New Zealand", "Singapore")

dat$Region <- factor(dat$Region,
                     levels = c("EU W. Europe",
                                "Americas",
                                "Asia Pacific",
                                "East EU Cemt Asia",
                                "MENA",
                                "SSA"),
                     labels = c("OECD",
                                "Americas",
                                "Asia &\nOceania",
                                "Central &\nEastern Europe",
                                "Middle East &\nnorth Africa",
                                "Sub-Saharan\nAfrica"))

ggplot(dat, aes(x=CPI, y=HDI)) +
  theme(legend.position = "top", text=element_text(size=8)) +
  geom_smooth(mapping=aes(linetype='R2'),
              method='lm', formula=y~x+log(x), se=FALSE, 
              color='#DA3C2A', size=0.7) + 
  geom_point(aes(color=Region), shape=21, size=2, stroke=1.5) +
  # labelling points
  geom_text_repel(aes(label=Country),
                  color='grey20',
                  data=subset(dat, Country %in% pointsToLabel),
                  size=2.5,
                  force=10) +
  
  scale_x_continuous(name="Human Development Index, 2011 (1=best)", limits = c(0.9,10.5), breaks = 1:10) +
  scale_y_continuous(name="Corruption Perceptions Index, 2011 (10=least corrupt)", limits = c(0.2,1.0), breaks = seq(0.1, 1, by=0.1)) + 
  scale_color_manual(name="", values = c("#24576D","#099DD7","#28AADC","#248E84","#F2583F","#96503F"), guide=guide_legend(nrow = 1)) +
  ggtitle("Corruption and Human Development")
  # + theme_bw()
```


##### Graph Simulation - Solution
```{r}
mR2 <- summary(lm(HDI ~ CPI + log(CPI), data = dat))$r.squared
mR2 <- paste0(format(mR2, digits = 2), "%")
```


```{r}
library(ggthemes)
library(ggrepel)

ggplot(dat, aes(x=CPI, y=HDI)) +
  geom_smooth(mapping=aes(linetype="r2"), method='lm', formula=y~x+log(x), se=FALSE, color='#DA3C2A', size=0.8) +
  geom_point(aes(color=Region), shape=1, stroke=1.5) +
  geom_text_repel(mapping=aes(label=Country, alpha=labels),
                  data=transform(dat, labels=Country %in% c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan",
                                                           "Afghanistan", "Congo", "Greece", "Argentina", "Brazil",
                                                           "India", "Italy", "China", "South Africa", "Spane",
                                                           "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France",
                                                           "United States", "Germany", "Britain", "Barbados", "Norway", "Japan",
                                                           "New Zealand", "Singapore")),
                  color='gray40',
                  segment.color='gray80',
                  size=3) +
  scale_alpha_discrete(range=c(0,1), guide=FALSE) +
  scale_x_continuous(name="Corruption Preception Index, 2011 (10=least corrupt)",
                     limits=c(1.0, 10.0),
                     breaks=1:10) +
  scale_y_continuous(name="Human Development Index, 2011 (1=best)",
                     limits=c(0.2, 1.0),
                     breaks=seq(0.2,1,by=0.1)) +
  scale_color_manual(name="",
                     values=c("#24576D","#099DD7","#28AADC","#248E84","#F2583F","#96503F"),
                     guide=guide_legend(nrow=1)) + 
  scale_linetype(name='',
                 breaks='r2',
                 labels=list(bquote(R^2==.(mR2))),
                 guide=guide_legend(override.aes = list(linetype=1, size=2, color="#DA3C2A"))) +
  ggtitle("Corruption and human development") +
  
  theme_bw() +
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        panel.grid.major.y = element_line(color='gray'),
        #panel.grid.major.x = element_line(color='gray'),
        
        axis.line.x = element_line(color='gray'),
        axis.text = element_text(face='italic'),
        axis.title.x = element_text(face='italic', size=8, color='gray20'),
        axis.title.y = element_text(face='italic', size=8, color='gray20'),
        
        legend.position = 'top',
        legend.direction = 'horizontal',
        legend.box= 'horizontal',
        legend.text = element_text(size=8,color='gray20'),
        
        plot.title = element_text(size=13, face = 'bold')
        )
  
```



http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization
















